---
title: "Richard Leyshon: Schools Analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: simplex
---


```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readxl)
library(purrr)
library(broom)
library(tibble)
library(stringr)
library(plotly)
library(knitr)
library(DT)
library(stringi)
library(leaflet)
library(leaflet.extras)
library(stringr)
library(mapview)
library(htmlwidgets)
library(ggmap)
```



```{r import, options}

schoolsxl <- read_excel("DelegatedSchoolsBudget.xlsx",
           skip = 3)

schoolsxl <- schoolsxl[, -1]
names(schoolsxl)[1] <- "Local_Authority"
names(schoolsxl)[2] <- "School"
names(schoolsxl)[3] <- "Pupil_Numbers"
names(schoolsxl)[4] <- "ISB(£_Thousands)"
schoolsxl <- schoolsxl[, -c(5:7)]


```


```{r remove NAs, options}

schoolsNArem <- schoolsxl %>% 
  separate(Local_Authority, sep = " - ", into = c("Local_Authority", "Phase"), remove = TRUE) %>% 
  fill(Local_Authority) %>% 
  fill(Phase) %>% 
  mutate(Pupil_Numbers = as.numeric(Pupil_Numbers),
         `ISB(Â£_Thousands)` = as.numeric(`ISB(Â£_Thousands)`)) %>% 
  filter(!is.na(Pupil_Numbers)) %>% 
  filter(!is.na(`ISB(Â£_Thousands)`)) %>%
  filter(!is.na(School)) 

```



```{r facet Phase, options}
#cleansing carried out in previous markdown, plus creation of composite key, LEA name and school number

schools_cleansed <- schoolsNArem %>% 
  separate(School, sep = 7, into = c("School Code", "schoolname"), remove = TRUE) %>% 
  type.convert("School Code")
  
schools_cleansed$schoolname <- sub('-', '', schools_cleansed$schoolname) 

schools_cleansed <- schools_cleansed %>%
  mutate(schoolname = str_squish(schoolname)) %>% 
  mutate(`School Code` = stri_sub(from = 4, to = 7, `School Code`)) %>%
  mutate('composite key' = paste(Local_Authority, `School Code`))

```



```{r prim_cleansed, options}

cleansedprimaries <- schools_cleansed %>%
  mutate(Phase = str_squish(Phase)) %>%
  filter(Phase == "Primary")

```

```{r middle_cleansed, options}
cleansedmiddles <- schools_cleansed %>%
  mutate(Phase = str_squish(Phase)) %>%
  filter(Phase == "Middle")

```


```{r spec_cleansed, options}
cleansedspecial <- schools_cleansed %>%
  mutate(Phase = str_squish(Phase)) %>%
  filter(Phase == "Special")

```


```{r sec_cleansed, options}

cleansedsecondary <- schools_cleansed %>%
  mutate(Phase = str_squish(Phase)) %>%
  filter(Phase == "Secondary")

```

```{r nestedandtidymodel, options}
# Nest phases to fit linear models by phase
phase_nested <- schools_cleansed %>%
  nest(-Phase)

#map linear model over nested phases, use mutate to store in column 'model'
nested_model <- phase_nested %>%
  mutate(model = map(data, ~lm(`ISB(Â£_Thousands)` ~ Pupil_Numbers, .)))

tidied_model <- nested_model %>%
  mutate(tidied = map(model, tidy))

unnested_tidied <- tidied_model %>%
  unnest(tidied)


phase_padjusted <- unnested_tidied %>%
  mutate(p.adjusted = p.adjust(p.value))


```



```{r v1.3, options}
#Import geocoded addresses created in schools dashboard v1.2
full_addresses <- read.csv("school_geocode.csv")

#Import delegated education dataset this time with all columns
#cnames_1.3 <- c("Local.Authority", "School", "Pupil numbers", "ISB (Â£ Thousand)", "ISB Per Pupil (Â£)", "Notional SEN Budget (Â£ Thousand)", "Non-ISB Funds Devolved to Schools (Â£ Thousand)")

#ctype_1.3 <- c("skip", "text", "text", "numeric", "numeric", "numeric", "numeric", "numeric")

schoolsxl_1.3 <- read_excel("DelegatedSchoolsBudget.xlsx",
#           col_names = cnames_1.3,
 #          col_types = ctype_1.3,
           skip = 3
           )

names(schoolsxl_1.3)[2] <- "Local.Authority"
names(schoolsxl_1.3)[3] <- "School"
names(schoolsxl_1.3)[4] <- "Pupil numbers"
names(schoolsxl_1.3)[5] <- "ISB (Â£ Thousand)"
names(schoolsxl_1.3)[6] <- "ISB Per Pupil (Â£)"
names(schoolsxl_1.3)[7] <- "Notional SEN Budget (Â£ Thousand)"
names(schoolsxl_1.3)[8] <- "Non-ISB Funds Devolved to Schools (Â£ Thousand)"
schoolsxl_1.3[1] <- NULL


#Perform the cleansing steps once more
schoolsNArem_1.3 <- schoolsxl_1.3 %>% 
  separate(`Local.Authority`, sep = " - ", into = c("Local.Authority", "Phase"), remove = TRUE) %>% 
  fill(`Local.Authority`) %>% 
  fill(Phase) %>% 
  filter(!is.na(`Pupil numbers`)) %>% 
  filter(!is.na("ISB (Â£ Thousand)")) %>%
  filter(!is.na(School))

schools_cleansed_1.3 <- schoolsNArem_1.3 %>% 
  separate(School, sep = 7, into = c("School Code", "schoolname"), remove = TRUE) %>% 
  type.convert("School Code")
  
schools_cleansed_1.3$schoolname <- sub('-', '', schools_cleansed_1.3$schoolname) 

schools_cleansed_1.3 <- schools_cleansed_1.3 %>%
  mutate(schoolname = str_squish(schoolname)) %>% 
  mutate(`School Code` = stri_sub(from = 4, to = 7, `School Code`)) %>%
  mutate('composite key' = paste(`Local.Authority`, `School Code`))

```

```{r new_join_1.3, options}

#rename key in full addresses
full_addresses4join <- full_addresses %>% 
  mutate(School.Code = stri_sub(from = 4, to = 7, School.Number)) %>%
  mutate(composite.key = paste(Local.Authority, School.Code))
 # rename(`composite key` = composite.key)

#Perform join based on composite key
joined_schools_1.3 <- left_join(full_addresses4join, schools_cleansed_1.3, by = c("composite.key" = "composite key"))



```


```{r TidyUp, options}
#Remove duplicate columns
map_dataset <- joined_schools_1.3[, -c(1, 6, 7, 9:14, 17, 20:22, 25)]

#User Friendly Names
cleansed_map_dataset <- map_dataset %>%
  rename(
    "Local.Authority" = Local.Authority.x,
    "Language Medium" = Welsh.Medium.Type...see.notes,
    "Phone Number" = Phone.Number
  )
cleansed_map_dataset$lat <- jitter(cleansed_map_dataset$lat, amount = 0.001)
cleansed_map_dataset$lon <- jitter(cleansed_map_dataset$lon, amount = 0.001)

```

Interactive Map
=========================================================================

### 2018/19 Individual Schools Budget Analysis. Data Source = [Stats Wales](https://statswales.gov.wales/Catalogue/Local-Government/Finance/Revenue/Delegated-School-Budgets/delegatedschoolbudgets-by-authority)
*Please note that a small amound of random noise has been added to school location coordinates in order to avoid overplotting schools that share a campus*

```{r label, options}

joined_nurseries <- cleansed_map_dataset %>% 
  filter(Phase == "Nursery")

joined_primaries <- cleansed_map_dataset %>% 
  filter(Phase == "Primary")

joined_secondaries <- cleansed_map_dataset %>% 
  filter(Phase == "Secondary")

joined_middles <- cleansed_map_dataset %>% 
  filter(Phase == "Middle")

joined_specials <- cleansed_map_dataset %>% 
  filter(Phase == "Special")





interactive_map <- leaflet() %>%
  addTiles(group = "OSM") %>% 
  addProviderTiles("Stamen.TonerLite", group = "Stamen") %>%
  addProviderTiles("Esri", group = "Esri") %>% 
  addCircleMarkers(data = joined_nurseries, group = "Nursery", color = "#004949",
                   popup = ~paste0("<b>", `School.Name`, "</b>", "<br>",
                                   "Pupil Numbers = ", `Pupil numbers`, "<br>",
                                   "ISB (Â£ Thousand) = ", "ISB (Â£ Thousand)", "<br>",
                                   "ISB Per Pupil (Â£) = ", `ISB Per Pupil (Â£)`, "<br>",
                                   "Notional SEN Budget (Â£ Thousand) = ", `Notional SEN Budget (Â£ Thousand)`, "<br>",
                                   "Non-ISB Funds Devolved to Schools (Â£ Thousand) = ", `Non-ISB Funds Devolved to Schools (Â£ Thousand)`, "<br>",
                                   `Local.Authority`),
                   opacity = 0.2) %>% 
  addCircleMarkers(data = joined_primaries, group = "Primary", color = "#ff6db6",
                   popup = ~paste0("<b>", `School.Name`, "</b>", "<br>",
                                   "Pupil Numbers = ", `Pupil numbers`, "<br>",
                                   "ISB (Â£ Thousand) = ", "ISB (Â£ Thousand)", "<br>",
                                   "ISB Per Pupil (Â£) = ", `ISB Per Pupil (Â£)`, "<br>",
                                   "Notional SEN Budget (Â£ Thousand) = ", `Notional SEN Budget (Â£ Thousand)`, "<br>",
                                   "Non-ISB Funds Devolved to Schools (Â£ Thousand) = ", `Non-ISB Funds Devolved to Schools (Â£ Thousand)`, "<br>",
                                   `Local.Authority`),
                   opacity = 0.2) %>% 
  addCircleMarkers(data = joined_secondaries, group = "Secondary", color = "#b66dff",
                   popup = ~paste0("<b>", `School.Name`, "</b>", "<br>",
                                   "Pupil Numbers = ", `Pupil numbers`, "<br>",
                                   "ISB (Â£ Thousand) = ", "ISB (Â£ Thousand)", "<br>",
                                   "ISB Per Pupil (Â£) = ", `ISB Per Pupil (Â£)`, "<br>",
                                   "Notional SEN Budget (Â£ Thousand) = ", `Notional SEN Budget (Â£ Thousand)`, "<br>",
                                   "Non-ISB Funds Devolved to Schools (Â£ Thousand) = ", `Non-ISB Funds Devolved to Schools (Â£ Thousand)`, "<br>",
                                   `Local.Authority`),
                   opacity = 0.2) %>%
  addCircleMarkers(data = joined_middles, group = "Middle", color = "#b6dbff",
                   popup = ~paste0("<b>", `School.Name`, "</b>", "<br>",
                                   "Pupil Numbers = ", `Pupil numbers`, "<br>",
                                   "ISB (Â£ Thousand) = ", "ISB (Â£ Thousand)", "<br>",
                                   "ISB Per Pupil (Â£) = ", `ISB Per Pupil (Â£)`, "<br>",
                                   "Notional SEN Budget (Â£ Thousand) = ", `Notional SEN Budget (Â£ Thousand)`, "<br>",
                                   "Non-ISB Funds Devolved to Schools (Â£ Thousand) = ", `Non-ISB Funds Devolved to Schools (Â£ Thousand)`, "<br>",
                                   `Local.Authority`),
                   opacity = 0.8) %>%
  addCircleMarkers(data = joined_specials, group = "Special", color = "#24ff24",
                   popup = ~paste0("<b>", `School.Name`, "</b>", "<br>",
                                   "Pupil Numbers = ", `Pupil numbers`, "<br>",
                                   "ISB (Â£ Thousand) = ", "ISB (Â£ Thousand)", "<br>",
                                   "ISB Per Pupil (Â£) = ", `ISB Per Pupil (Â£)`, "<br>",
                                   "Notional SEN Budget (Â£ Thousand) = ", `Notional SEN Budget (Â£ Thousand)`, "<br>",
                                   "Non-ISB Funds Devolved to Schools (Â£ Thousand) = ", `Non-ISB Funds Devolved to Schools (Â£ Thousand)`, "<br>",
                                   `Local.Authority`),
                   opacity = 0.2) %>% 
  addLayersControl(
    options = layersControlOptions(collapsed = FALSE),
    baseGroups = c("Stamen", "OSM", "Esri"), 
    overlayGroups = c("Nursery", "Primary", "Secondary", "Middle", "Special")
  ) %>%
  addResetMapButton() %>% 
  addLegend(position = "bottomright",
            labels = c("Nursery", "Primary", "Secondary", "Middle", "Special"),
            colors = c("#004949", "#ff6db6", "#b66dff", "#b6dbff", "#24ff24")) %>% 
  addMouseCoordinates() 


interactive_map
```




Primaries {.storyboard}
=======================================================================



### Primary Phase: Individual Schools Budget vs Pupil Numbers

```{r primary_phaseplot}

primary_static <- cleansedprimaries %>%
  ggplot(aes(x = Pupil_Numbers, y = `ISB(Â£_Thousands)`, label = schoolname)) +
  geom_point(alpha = 0.3) +
  geom_smooth()

ggplotly(primary_static)
 
```



### Linear Regression Analysis

```{r Prim_Model, options}

primary_model <- phase_padjusted %>% 
  filter(Phase == "Primary")
primary_model


```
The regression shows that for the primary phase in Welsh schools, each individual student increases the Individual Schools Budget by Â£3,200. The p value returns a strong statistical significance (p adjusted = 0). Essentially, in 99.9% of cases, an increase in ISB is explained by an increase in pupil numbers. The standard error is extremely low, at a value of 0.0247. This means that on average, the model will be incorrect by Â£25 plus or minus the median value. 

Note that the nursery phase returned a non-significant p value and therefore was not analysed. This is likely due to the small number of nursery schools within Wales. 


### Interactive Data Table

```{r prim_datatable, options}
cleansedprimaries %>% 
  datatable(rownames = F,
            extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
            ))

```

Secondaries {.storyboard}
======================================================================================

### Secondary Phase: Individual Schools Budget vs Pupil Numbers

```{r secondary_plotly, options}
secondary_static <- cleansedsecondary %>%
  ggplot(aes(x = Pupil_Numbers, y = `ISB(Â£_Thousands)`, label = schoolname)) +
  geom_point(alpha = 0.3) +
  geom_smooth()

ggplotly(secondary_static)

```

### Linear Regression Analysis

```{r Sec_Model, options}
secondary_model <- phase_padjusted %>% 
  filter(Phase == "Secondary")
secondary_model

```
The regression shows that for the secondary phase in Welsh schools, each individual student increases the Individual Schools Budget by Â£4,150. The p value returns a strong statistical significance (p adjusted = 1.38e-126). In 99.9% of cases, an increase in ISB is explained by an increase in pupil numbers. The standard error is extremely low for this phase, at a value of 0.0701. This means that on average, the model will be incorrect by Â£70 plus or minus the median value. 

### Interactive Data Table
```{r secondarydatatable, options}
cleansedsecondary %>% 
  datatable(rownames = F,
            extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
            ))

```

Middles {.storyboard}
================================================================================


### Middle Phase: Individual Schools Budget vs Pupil Numbers

```{r middle_plotly}

middle_static <- cleansedmiddles %>% 
  ggplot(aes(x = Pupil_Numbers, y = `ISB(Â£_Thousands)`, label = schoolname)) +
  geom_point(alpha = 0.3) +
  geom_smooth()

ggplotly(middle_static)
```



### Linear Regression Analysis

```{r Mid_Model, options}
middle_model <- phase_padjusted %>% 
  filter(Phase == "Middle")
middle_model

```

The regression shows that for the middle phase in Welsh schools, each individual student increases the Individual Schools Budget by Â£4,390. The p value returns a strong statistical significance (p adjusted = 2.53e-12). In 99.9% of cases, an increase in ISB is explained by an increase in pupil numbers. The standard error is very low, at a value of 0.24. This means that on average, the model will be incorrect by Â£224 plus or minus the median value. 

### Interactive Data Table

```{r middledatatable, options}
cleansedmiddles %>% 
  datatable(rownames = F,
            extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
            ))

```

Specials {.storyboard}
========================================================================================


### Special Phase: Individual Schools Budget vs Pupil Numbers

```{r Special_plotly}

special_static <- cleansedspecial %>%
  ggplot(aes(x = Pupil_Numbers, y = `ISB(Â£_Thousands)`, label = schoolname)) +
  geom_point(alpha = 0.3) +
  geom_smooth()

ggplotly(special_static)

```


### Linear Regression Analysis

```{r Spec_Model, options}
special_model <- phase_padjusted %>% 
  filter(Phase == "Special")
special_model

```
The regression shows that for the special phase in Welsh schools, each individual student increases the Individual Schools Budget by Â£18,300. The p value returns a strong statistical significance (p adjusted = 6.63e-13). In 99.9% of cases, an increase in ISB is explained by an increase in pupil numbers. 

However, the standard error is relatively larger than the other phases in this report, at a value of 1.63. This means that on average, the model will be incorrect by Â£1,630 plus or minus the median value. This increase in standard error is symptomatic of having a relatively small sample size in relation to the correlation coefficient and therefore would advise caution in drawing conclusions from this dataset.

### Interactive Data Table
```{r specialdatatable, options}
cleansedspecial %>% 
  datatable(rownames = F,
            extensions = 'Buttons',
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
            ))

```