---
title: "Unnamed"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---
**This markdown is broken. Your job is to get it up and running again.\n** *Delete once understood.*

I've made lots of errors throughout. I keep getting Python code mixed up with R code. The good news is that I remembered to flag these issues. Wherever these problems occur, I've labelled them with 'task'. If you ctrl + F you can find all those instances. I would prioritise the issues you find within the code chunks first, from there on you should be able to knit the dashboard. *Delete once understood.*


**task: ** edit the title in the YAML to something better & delete this line. *Delete once actioned.*

```{r setup, include=FALSE}
library(flexdashboard)
#dplyr, mainly for the pipes. Hide the conflict messages. 
library(dplyr, warn.conflicts = FALSE)
#ggplot2, for the charts
library(ggplot2, warn.conflicts = FALSE)
#plotly for interactive charts
library(plotly, warn.conflicts = FALSE)
#DT for interactive data table
library(DT, warn.conflicts = FALSE)

#load the MT cars dataset
cars_data <- mtcars
```



**task: ** Add 2 pages:

Page 1: Title - 'Exploratory Data Analysis'.
Contains the first 3 charts (code chunks).

Page 2: Title - 'See the data'.
Contains the last code chunk with the data table.

For help on how to do this, follow [this link](https://rmarkdown.rstudio.com/flexdashboard/)


Column 
-----------------------------------------------------------------------

**task: ** this column width should be set to 650 px *Delete once actioned.*

### How does vehicle weight affect fuel efficiency?

```{r, echo = FALSE}
# In this code chunk, let's start with the good old scatter chart. But we'll make it interactive. I thought it may be interesting to investigate how the weight of a vehicle affects its fuel efficiency.

#create a static ggplot that we can pass to ggplotly later for simple interactivity
static_plot <- cars_data %>%
  #pipe our cars data into ggplot
  ggplot(
    #set the aesthetics for the dimensions of the chart
    aes(
      #vehicle weight on x axis
      x = wt,
      #vehicle mpg on y axis
      y = mpg,
      #add a text aesthetic. Plotly can render this as a hover tooltip. 
      text = paste("model:", row.names(cars_data))
      )#end of aes
    ) + #end of ggplot base layer
  
  #task: What's up with my scatterplot?
  #add point data to the axes
  sns.scatterplot() + 
  #edit the charts axes labels
  #task: I've messed up the labels somehow
  labs(
    x = "Mass of vehicle (1000 lbs)",
    plt.ylabel("Fuel efficiency (MPG)")
  ) + #end of labels
  #apply a theme to remove the grey back panel from the ggplot
  theme_bw()
  
#pass the static ggplot to the plotly::ggplotly function adds some simple 'out of the box' interactivity. Note that plotly has its own syntax for building charts, too. This is a bit more rigorous than the method we are using here. 
static_plot %>% ggplotly()

```

Column 
-----------------------------------------------------------------------

**task: ** this column width should be set to 350 px *Delete once actioned.*

### Fuel efficiencies for the top 7 most powerful vehicles

```{r}

# In this code chunk, I've gone for a static plot. I couldn't see how interactivity would add value here, so it would just be a distraction. 
# 
# I wanted to look at the fuel efficiency for the most powerful cars in the dataset. This time a bar chart of the top 7 powerful cars. I'd like the model names on the x axis & horsepower on the y axis. To colour the bars by mpg, you need to use the fill aesthetic (colour would adjust the bar outlines). 

#curate the data. Just the top 7 powerful vehicles for this task.
top7Powerful <- cars_data %>% 
  #top_n will slice the dataframe. Give it the number of rows you want and the column to rank on
  top_n(7, hp)
  
#As we're not going for interactivity, no need to assign the chart to an object
top7Powerful %>% 
  #create the ggplot base layer
  ggplot(
    #add data dimensions to the chart aesthetics
    aes(
      #the vehicle models are stored in the dataframe rownames (like index in python).
      #ggplot tries to plot the models alphabetically. To reorder by power, use base R's
      #reorder()
      x = reorder(rownames(top7Powerful), hp),
      #power on the y axis
      y = hp,
      #fuel efficiency to the fill aesthetic
      fill = mpg
      )#end of aesthetics
    ) + #end of ggplot base layer
  #add bars to the base layer
  geom_col() +
  #the vehicle names are too crowded, coord_flip realigns the graphic horizontally 
  coord_flip() +
  #remove the greay back panel again
  theme_bw() +
  #customise the axes labels
  #task: again with the labels
  labs(
    plt.xlabel("Vehicle Model"),
    y = "Horsepower"
    )#end of labels


```

### How does the number of gears affect fuel efficiency distribution?

```{r}

# In this chart I wanted to investigate whether cars with more gears had better fuel economy. I wanted to see the distribution of fuel efficiencies for the vehicles. Therefore I've chosen a boxplot. As the dataset is American in origin, there are mostly automatic transmission cars (am == 0). I chose to highlight this in the chart by overlaying the points too, and colouting them based on their transmission type.

#task: no syntax errors in this code chunk, but I've managed to muddle my ordering somehow...
  
  



#pass to plotly::ggplotly() for interactivity, constrain the tooltip to just the y aesthetic
ggplotly(static_barchart, tooltip = 'y')



#this one will be interactive, so let's assign the ggplot to a variable
static_barchart <- cars_data %>% 


  #add the boxplot layer
  geom_boxplot(
    #set the fill colour
    fill = 'lightblue',
    #adjust the opacity to allow contrast with the points
    alpha = 0.4
    ) + #end of boxplot
  #also plot points but with a small degree of random noise.(Helps with overplotting)
  geom_jitter(
    #I want to relabel any row where am column is 0 to auto, else label it manual. Map this to
    #colour.
    aes(colour = ifelse(am == 0, "automatic", "manual")),
    #I don't want any random noise on the mpg scale
    height = 0,
    #just a small amound of noise on the gear scale to separate clustered points
    width = 0.01,
    #adjusting transparency helps to avoid overplotting issues
    alpha = 0.4
    ) + #end of geom_jitter
  #remove the grey panel
  theme_bw() + 
  #label the axes & legend
  labs(
    x = "Number of forward gears",
    y = "Fuel efficiency (mpg)",
    colour = "Transmission"
    )


  #add the ggplot base layer
  ggplot(
    #map dimensions to aesthetics
    aes(
      #gear data type is numeric. make it discrete by coercing to factor. 
      x =  as.factor(gear),
      #fuel efficiency on y axis
      y = mpg) #end of aes
    ) + #end of ggplot base layer







```



Column 
-----------------------------------------------------------------------

**task: ** specify the column width to 1000 px

### Interactive Data Table



```{r}

#pipe cars_data into the datatable
cars_data %>% datatable(
  #rename all columns to make them user friendly, use c() to combine a vector of comma 
  #separated values
  
  
#task: can you present the column names of the data table in a more user-friendly way?
#See https://rstudio.github.io/DT/ for help here.  
  
  
  
#Add some export options Use [this link](https://rstudio.github.io/DT/003-tabletools-buttons.html) to help.   
  
  #add the buttons extensions to the datatable

  #edit the button options, use a list for all required options

    #dom sets where the buttons will be located

    #assign the required button options to a c() combined vector of comma separated values

      #copy to clipboard

      #export to csv

      #export to xlsx

      #export to pdf

      #print to OneNote / printer


  
)#end of datatable

```

