---
title: "Dependencies for Basic Stats in DS"
author: "R Leyshon"
date: "20/11/2020"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
install.packages("reticulate", type = "win.binary", dependencies = TRUE)
library(reticulate)

```



# Dependencies

This course uses the `reticulate` package to run python code chunks from R markdown. In order to run all chunks and knit the markdown, you will need to have python installed on your computer. You may also need to run the following code if you are receiving errors that look like this: `Error in path.expand(path) : invalid 'path' argument`.

```{r}
# library(reticulate)
# use_python("/Library/Frameworks/Python.framework/Versions/3.7/bin/python3")
# py_run_string("import os as os")
# py_run_string("os.environ['QT_QPA_PLATFORM_PLUGIN_PATH'] = 'C:/Python36/Library/plugins/platforms'")

```

Ensure that you locate the path to your python executable and replace the path included in the `use_python` function above. On an ONS laptop, I found that reticulate worked fine out of the box when running R version 3.4.0, but that later versions would not run python code chunks.

Alternatively, you can try:

```{python}
import os
os.environ['QT_QPA_PLATFORM_PLUGIN_PATH'] = 'C:\Python36\Library\plugins\platforms'
# though you will need to ensure the path is correct for you
```


If everything has been configured correctly, you should be able to run the below code chunk, and **crucially** also be able to knit to html (either by using the knit button in the top menu or by pressing `ctrl` + `shift` + `k`).


```{python}
import matplotlib.pyplot as plt

import random
#Generate 5 random numbers between 10 and 30
randomlist = random.sample(range(10, 50), 25)
print(randomlist)


x = plt.hist(randomlist)

plt.show(x)

```

**If you can see the plot in the html output, you're good to run python chunks within the course.**

If the above does not resolve your situation, please let me know and I will look into these issues for further training development. If issues are experienced running python code chunks, I suggest running the alternative markdown which contains R exercises only.



